---
title: "Mapping with ggplot"
author: "Meddy Rutayisire"
format: html
---

Class was inspired by works by Mel Moreno and Mathieu Basile found [Here](https://r-spatial.org/r/2018/10/25/ggplot2-sf.html)

# Introduction to Mapping in R

## Overview

Welcome to the "Introduction to Mapping in R" class! This course is designed to provide you with the foundational skills needed to create and visualize geographic data using R.

R is a powerful statistical programming language that offers extensive libraries and tools for working with spatial data, making it an excellent choice for mapping and geographic information system (GIS) applications.

## Objectives

By the end of this course, you will be able to: - Understand the basics of spatial data and geographic coordinate systems.

Import and manipulate spatial data in R. Create a variety of maps using popular R packages. Customize maps to highlight specific data insights.

## Getting Started

```{r}
pacman::p_load(
 tidyverse,sf , rnaturalearth
)
```

### Set Theme

```{r}
theme_set(theme_bw())
```

The package rnaturalearth provides a map of countries of the entire world. Use ne_countries to pull country data and choose the scale (rnaturalearthhires is necessary for scale = "large"). The function can return sp classes (default) or directly sf classes, as defined in the argument returnclass:

```{r}
world_df <- ne_countries(scale = "medium", returnclass = "sf")

class(world_df)
sample(names(world_df),10)
```

### Data and basic plot (ggplot and geom_sf)

```{r}
ggplot(data = world_df) +
    geom_sf()
```

### Add labels

```{r}
ggplot(data = world_df) +
    geom_sf() +
    xlab("Longitude") + ylab("Latitude") +
    labs(title = "World map")
```

### Add Color

```{r}
ggplot(data = world_df) + 
    geom_sf(color = "black", fill = "lightgreen")
```

```{r}
ggplot(data = world_df) +
    geom_sf(aes(fill = pop_est))
```

