---
title: "Malaria Report"
format: pdf
---

## Summary Table

```{r}
pacman::p_load(
  readxl, ## read xlsx
  tidyverse, ## several packages for wrangling 
  here,  ## locate files , 
  gt
)

## load data 

data_malaria <- read_excel(here("District_data.xlsx"))

```

## Districts with the most Deaths 
```{r}
top_districts_deaths <- data_malaria |> 
  filter(grepl("No. of deaths", dataelement)) |> 
  group_by(district) |> 
  summarise(deaths = sum(value,na.rm = T)) |> 
  slice_max(order_by = deaths , n = 10)

top_districts_deaths |> 
  gt()
  
```


## Malaria Cases Per 1,000 population

### Monthly trend (national)

### Map showing Distribution by District (Last 4 years)

## Malaria Testing Rate

### Monthly trend (national)

### Map showing Distribution by District (Last 4 years)

### Point map with point size at facility level (Last 4 years)

## Malaria Test Positivity Rate

### Monthly trend (national)

### Map showing Distribution by District (Last 4 years)

## Malaria as percentage of Total OPD

### Monthly trend (national)

### Map showing Distribution by District (Last 4 years)

### Malaria deaths per 100,000 population

### Monthly trend (national)

### Map showing Distribution by District (Last 4 years)

## Malaria deaths as % of Total Deaths

### Monthly trend (national)

### Map showing Distribution by District (Last 4 years)

## Malaria case fatality Rate

### Monthly trend (national)

### Map showing Distribution by District (Last 4 years)
